---
title: "R day 2"
author: "Anjali Verma"
date: "9/18/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
here::here("data", "gapminder5.csv")
gapminder<-read.csv(here::here("data","gapminder5.csv"))
```
## Converting factors to characters
```{r}
gapminder$country <- as.character(gapminder$country)
gapminder$continent <- as.character(gapminder$continent)
str(gapminder)
```
## Creating logged variables with a for loop
```{r}
conts <- unique(gapminder$continent)
for(i in conts){mean_le <- mean(gapminder$lifeExp[gapminder$continent == i], na.rm = TRUE)
 print(paste0(i ,":", mean_le))}

```
```{r}
conts <- unique(gapminder$continent)
years <- unique(gapminder$year)
for (i in conts){print(paste0("Continent",": ", i))
for(j in years){sd_le <- sd(gapminder$lifeExp[gapminder$continent == i & gapminder$year == j],na.rm = TRUE)
print(paste0(j ,":", sd_le))}}
```
```{r}
i <- 1952
while(i < 1987){
  sd_le <-sd(gapminder$lifeExp[gapminder$year == i], na.rm = TRUE)
  
  print(paste0(i, ":", sd_le))
  i <- i+5
}
```
```{r}

random_year <- sample(years,1)
if (random_year>1977){print(random_year)}else{
  print("sorry, random year is less than 1977")
}
```
```{r}
for (i in unique(gapminder$year)){
  if(i >= 1987){
    mean_pop <- mean(gapminder$pop[gapminder$year == i], na.rm =TRUE)
    print(paste0(i, ":", mean_pop))
  }else{print("Sorry, the year is less than 1987")}
}
```
```{r}
statistics_gapminder <- function(df,variable,continent){
  var <- df[[variable]][df$continent == continent]
  mean_le <- mean(var)
  sd_le <- sd(var)
  max_le <- max(var)
  min_le <- min(var)
  cat("Continent:", continent, 
        "\nMean Life Expectancy:", mean_le,
        "\nSD Life Expectancy:", sd_le,
      "\nMax Life Expectancy:", max_le,
      "\nMin Life Expectancy:", min_le)
}
statistics_gapminder(gapminder,"lifeExp","Asia")
```
## Load the data
```{r}
here::here("data","gapminder5.csv")
gapminder <- read.csv(here::here("data","gapminder5.csv"))
```
## If Statement
```{r}
if(any(gapminder$year == 2002)){print("Information on 2002 is present")
}else{print("Information on 2002 is not present")}
if(any(gapminder$year == 2012)){print("Information on 2012 is present")
  }else{print("Information on 2012 is not present")}
```
## Loop and If statements
```{r}
threshold <- mean(gapminder$pop)
for(i in unique(gapminder$country)){
  if(mean(gapminder$pop[gapminder$country == i])<threshold){
    mean_le<- mean(gapminder$lifeExp[gapminder$country == i])
    print(paste0(i,":", mean_le))
  }else{
    print("Mean population for country is greater than threshold ")
  }
}
```
```{r}

for(i in unique(gapminder$continent)){
  if(mean(gapminder$lifeExp[gapminder$continent == i]) < 50){print(paste0(i,":","Less than 50"))}
  if(mean(gapminder$lifeExp[gapminder$continent == i]) > 70){print(paste0(i,":","Greater than 70"))}
  else{print(paste0(i,":","Greater than 50 and less than 70"))}
}

```
## Writing Functions
```{r}
class_and_names <- function(df){
for(i in 1:ncol(df))
{colnames<-colnames(df)
 print(paste(colnames[i],":",class(df[,i])))
}
}
class_and_names(gapminder)

```
```{r}
vector_statistics<- function(vector, option = FALSE){
  mean <- mean(vector)
  median<- median(vector)
  sd <- sd(vector)
  if(option == TRUE){
    print(paste0(mean," ", sd," ", median))
  }else{
    print(paste0(mean," ", sd))
  }
}
vector_statistics(unique(gapminder$year), option = TRUE)
```
## Analyzing Relationships
```{r}
viz_lm <- function(df, dv, iv, year){
dat <- df[df[["year"]] == year,]
y <- log(dat[[dv]])
x <- log(dat[[iv]])
fit <- lm(y~x)
plot(y~x, main = year, xlab = iv, ylab = dv)
lines(x, predict(fit), col = 'blue')
}
for (i in years){viz_lm(gapminder, "lifeExp","gdpPercap",i)}
```
## There is a positive relationship between lifeExp and gdpPercap that is, a higher GDP corresponds to a higher life expectancy. However, over the years, the line of fit becomes more flat which means that the smae change in GDP coresponds to a relatively lower change in life expectancy, as the years go by.

```{r}
viz_lm <- function(df, dv, iv, continent){
dat <- df[df[["continent"]] == continent,]
y <- log(dat[[dv]])
x <- log(dat[[iv]])
fit <- lm(y~x)
plot(y~x, main = continent, xlab = iv, ylab = dv)
lines(x, predict(fit), col = 'blue')
}
for (i in conts){viz_lm(gapminder, "lifeExp","gdpPercap",i)}

```